Precise predictions for same-sign W-boson scattering at the LHC

### 2018-03-04 Draft/contributions.tex
Start of implementing Ansgar's comments

% At tree level, there are three contributions to the $\PW^+\PW^+$ production in association with two jets: the pure EW component $\mathcal{O}(\alpha^6)$, the interference $\mathcal{O}(\alphas\alpha^5)$, and the QCD background $\mathcal{O}(\alphas^2\alpha^4)$
% \AK{We have mentioned this many times at this point. Perhaps not worth repeating again?}.
In the present section, the cross sections and distributions are obtained without applying the VBS cuts on $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$, 
Eq.~(\ref{cut:4}).
In Tab.~\ref{tab:LOscanXsec}, the cross sections of the three contributions are reported.
The EW, QCD, and interference contributions amount to $57\%$, $37\%$, and $6\%$ of the total inclusive cross section, respectively.
The QCD contribution does not posses external gluons due to charge conservation.
Thus the diagrams of order $\mathcal{O}(\alphas^2\alpha^4)$ only involve gluon exchange in the $t/u$-channel between the quark lines.
This results in a small contribution although the VBS cuts have not been imposed.
The interference between EW and QCD contributions is small, due to color suppression, but not negligible.
% ($t/u$ and $t/s$ interferences with identical fermions).

@@

In Fig.~\ref{fig:mjjdyjj_1d} these three contributions are shown separately and summed in the differential distribution of the di-jet invariant mass $m_{\Pj\Pj}$ and the rapidity difference $|\Delta y_{\Pj\Pj}|$.
For the distribution in the di-jet invariant mass (left), one can observe that the EW contribution peaks around an invariant mass of about $80\GeV$.
These are due to diagrams where the two jets originate from the decay of a W boson (see middle and right diagrams in Fig.~\ref{diag:LO}).
Note that these contributions are not present in calculations relying on the VBS approximation.
The EW contribution becomes dominant for di-jet invariant mass larger than $500\GeV$.
The same holds true for jet rapidity difference larger than $2.5$ (right).
This justifies why cuts on these two observables are used in order to enhance the EW contribution over the QCD one.
In particular, in order to have a large EW contribution, rather exclusive cuts are required.

>>>>

% At tree level, there are three contributions to the $\PW^+\PW^+$ production in association with two jets: the pure EW component $\mathcal{O}(\alpha^6)$, the interference $\mathcal{O}(\alphas\alpha^5)$, and the QCD background $\mathcal{O}(\alphas^2\alpha^4)$
% \AK{We have mentioned this many times at this point. Perhaps not worth repeating again?}.
In the present section, the cross sections and distributions are obtained without applying the VBS cuts on $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$, 
Eq.~(\ref{cut:4}).
In Tab.~\ref{tab:LOscanXsec}, the cross sections of the three LO contributions are reported.
The EW, QCD, and interference contributions amount to $57\%$, $37\%$, and $6\%$ of the total inclusive cross section, respectively.
The QCD contribution does not posses external gluons due to charge conservation.
Thus the diagrams of order $\mathcal{O}(\alphas^2\alpha^4)$ only involve gluon exchange in the $t/u$-channel between the quark lines.
This results in a small contribution even if the VBS cuts have not been imposed.
The interference between EW and QCD contributions is small, due to color suppression, but not negligible.
% ($t/u$ and $t/s$ interferences with identical fermions).

@@

In Fig.~\ref{fig:mjjdyjj_1d} these three contributions are shown separately and summed in the differential distribution of the di-jet invariant mass $m_{\Pj\Pj}$ and the rapidity difference $|\Delta y_{\Pj\Pj}|$.
For the distribution in the di-jet invariant mass (left), one can observe that the EW contribution peaks around an invariant mass of about $80\GeV$.
This is due to diagrams where the two jets originate from the decay of a W boson (see middle and right diagrams in Fig.~\ref{diag:LO}).
Note that these contributions are not present in calculations relying on the VBS approximation.
The EW contribution becomes dominant for di-jet invariant mass larger than $500\GeV$.
The same holds true for jet rapidity difference larger than $2.5$ (right).
This justifies why cuts on these two observables are used in order to enhance the EW contribution over the QCD one.
In particular, in order to have a large EW contribution, rather exclusive cuts are required.

### 2018-03-04 Draft/definition.tex
Start of implementing Ansgar's comments

This process possesses three LO contributions of different orders.
At LO, this process can proceed via three different coupling-order combinations:
$\mathcal{O}{\left(\alpha^{6}\right)}$, $\mathcal{O}{\left(\alphas{2}\alpha^{4}\right)}$, and $\mathcal{O}{\left(\alphas\alpha^{5}\right)}$.
The first, commonly referred to as EW contribution or VBS,\footnote{The name VBS is used even though not all Feynman diagrams involve the scattering
of vector bosons.} receives the contributions from Feynman diagrams such as those in Fig.~\ref{diag:LO}:
in addition to genuine VBS contributions (left diagram), it also features $s$-channel contributions with non-resonant vector bosons 
(center diagram) or from
triple-boson production (right diagram).
Note that $s$-, $t$-, and $u$-channel contributions are defined according to the quark lines.
$s$-channel denotes all Feynman diagrams where the two initial-state partons are connected by a continuous fermion line.
$u$-channel refers to contributions with crossed fermion lines, which appears for identical quarks or anti-quarks in the final state
The $s$-channel contributions will play a particular role in the study of the various contributions in Sec.~\ref{subsec:contributions}.

@@

Finally, the third contribution is the interference of the two types of amplitudes described above.
It is non-zero only for those partonic subprocesses which involve only one quark family.
Such a contribution is typically small but not negligible for realistic experimental set-ups \cite{Biedermann:2017bss}.

>>>>

This process possesses three LO contributions of different orders.
At LO, this process can proceed via three different coupling-order combinations:
$\mathcal{O}{\left(\alpha^{6}\right)}$, $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$, and $\mathcal{O}{\left(\alphas\alpha^{5}\right)}$.
The first, commonly referred to as EW contribution or VBS,\footnote{The name VBS is used even though not all Feynman diagrams involve the scattering
of vector bosons.} receives the contributions from Feynman diagrams such as those in Fig.~\ref{diag:LO}:
in addition to genuine VBS contributions (left diagram), it also features $s$-channel contributions with non-resonant vector bosons 
(center diagram) or from
triple-boson production (right diagram).
Note that $s$-, $t$-, and $u$-channel contributions are defined according to the quark lines.
$s$-channel denotes all Feynman diagrams where the two initial-state partons are connected by a continuous fermion line.
$u$-channel refers to contributions with crossed fermion lines with respect to $t$-channel, which appears for identical quarks or anti-quarks in the final state
The $s$-channel contributions will play a particular role in the study of the various contributions in Sec.~\ref{subsec:contributions}.

@@

Finally, the third contribution is the interference of the two types of amplitudes described above.
It is non-zero only for those partonic subprocesses which involve identical quarks or antiquarks.
Such a contribution is typically small but not negligible for realistic experimental set-ups \cite{Biedermann:2017bss}.

### 2018-03-04 Draft/input.tex
Start of implementing Ansgar's comments

The hadronic scattering processes are simulated at the LHC with a center-of-mass energy $\sqrt s = 13 \TeV$.
The NNPDF~3.0 parton distribution functions~(PDFs)~\cite{Ball:2014uwa} with five massles flavour,\footnote{For the process considered, no bottom (anti-)quarks appear in the initial or final state at LO and NLO.} 
NLO-QCD evolution, and a strong coupling constant $\alphas\left( \MZ \right) = 0.118$ are employed.\footnote{The corresponding identifier {\tt lhaid} in the program LHAPDF6~\cite{Buckley:2014ana} is 260000.}\footnote{Note that the {\sc Powheg-Box} uses its own implementation of the two loop running for $\alpha_{\rm s}$.}
%Since this PDF set has no photonic density, photon-induced processes are not considered.
Initial-state collinear singularities are factorised according to the ${\overline{\rm MS}}$ scheme, consistently with what is done in NNPDF.

>>>>

The hadronic scattering processes are simulated at the LHC with a center-of-mass energy $\sqrt s = 13 \TeV$.
The NNPDF~3.0 parton distribution functions~(PDFs)~\cite{Ball:2014uwa} with five massles flavours,\footnote{For the process considered, no bottom (anti-)quarks appear in the initial or final state at LO and NLO.} 
NLO-QCD evolution, and a strong coupling constant $\alphas\left( \MZ \right) = 0.118$ are employed.\footnote{The corresponding identifier {\tt lhaid} in the program LHAPDF6~\cite{Buckley:2014ana} is 260000.}\footnote{Note that the {\sc Powheg-Box} uses its own implementation of the two loop running for $\alpha_{\rm s}$.}
%Since this PDF set has no photonic density, photon-induced processes are not considered.
Initial-state collinear singularities are factorised according to the ${\overline{\rm MS}}$ scheme, consistently with what is done in NNPDF.

### 2018-03-04 Draft/process.tex
Start of implementing Ansgar's comments

We now discuss the various approximations which are implemented in computer programs for the EW contribution at order $\mathcal{O}{\left(\alpha^{6}\right)}$.
Since we are mostly interested in 
the scattering of two $\PW$ gauge bosons, which includes the quartic gauge-boson vertex, may appear justified to approximate the 
full process by considering just those diagrams which contain the $2\rightarrow 2$ scattering process as a sub-part.
However, this set of contributions is not gauge invariant.
In order ensure gauge invariance, an on-shell projection of the incoming and outgoing $\PW$ bosons should be performed {\bf MZ CITE}.
Unfortunately, the former momenta are space-like and thus a simple on-shell projection is not possible.
Instead, one can keep the $\PW$ boson legs connected to the external quark line off-shell while the ones connected 
to the final-state leptons, which are already time-like, are put on-shell. \ZM{is this still gauge invariant? if so, one should change 
the sentence above ``In order to ensure..''}
Then the polarisation of the gauge boson is accommodated following for example the implementations of Refs.~\cite{Kuss:1995yv,Accomando:2006hq}.
Such an approximation is usually called effective vector-boson approximation (EVBA) \cite{Dawson:1984gx,Duncan:1985vj,Cahn:1983ip}.

@@

When considering the full one-loop amplitude of order $\mathcal{O}(g_{\rm s}^2 g^6)$ squared with the tree amplitude of order $\mathcal{O}(g^6)$, not only real QCD radiation but also QED ones have to be included in order cancel IR singularities.
But all IR singularities related to photon emissions are not cancelled by the above mentioned virtual corrections.
Another type of virtual corrections has to be incorporated, namely the one-loop amplitude of order $\mathcal{O}(g^8)$ interfered with the tree-level one of order $\mathcal{O}(g_{\rm s}^2 g^4)$.
Hence the full NLO corrections of order $\mathcal{O}{\left(\alphas \alpha^{6}\right)}$ consist not only of QCD-type corrections but also of EW ones \cite{Biedermann:2017bss}.
%Such predictions are provided by the combination {\sc MoCaNLO+Recola} as published in Ref.~\cite{Biedermann:2017bss}.

>>>>

We now discuss the various approximations which are implemented in computer programs for the EW contribution at order $\mathcal{O}{\left(\alpha^{6}\right)}$.
Since we are mostly interested in 
the scattering of two $\PW$ gauge bosons, which includes the quartic gauge-boson vertex, it may appear justified to approximate the 
full process by considering just those diagrams which contain the $2\rightarrow 2$ scattering process as a sub-part.
However, this set of contributions is not gauge invariant.
In order ensure gauge invariance, an on-shell projection of
the incoming and outgoing W bosons should be performed.
While this can be done in the usual way for the time-like
outgoing W bosons, the treatment of the space-like W bosons
emitted from the incoming quarks requires some care.
Following Refs.~\cite{Kuss:1995yv,Accomando:2006hq} these W-boson lines can be split,
the W bosons entering the scattering process can be projected
on-shell, and the emission of the W bosons from the quarks can be
described by vector-boson luminosities.
Such an approximation is usually called effective vector-boson
approximation (EVBA) \cite{Dawson:1984gx,Duncan:1985vj,Cahn:1983ip}.

@@

When considering the NLO corrections of order
$\mathcal{O}{\left(\alpha_{\rm s}\alpha^{6}\right)}$
to the full process, besides real and virtual QCD corrections
to the EW tree-level contribution of order
$\mathcal{O}{\left(\alpha^{6}\right)}$
also real and virtual EW corrections to the LO interference
of order $\mathcal{O}{\left(\alpha_{\rm s}\alpha^{5}\right)}$
have to be taken into account. Since some loop diagrams contribute
to both types of corrections, QCD and EW corrections cannot be
separated any more on the basis of Feynman diagrams, and the
cancellation of IR singularities requires to include all of them \cite{Biedermann:2017bss}.

### 2018-03-04 Draft/scanLO.tex
Start of implementing Ansgar's comments

In Fig.~\ref{fig:ratio2d_LO}, ratios of double-differential cross sections in the plane $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$ are shown.\footnote{In Fig.~\ref{fig:ratio2d_LO}, the level of the accuracy of the predictions in each bin is around a per mille.}
Two plots are displayed: the ratios of the $|t|^2 + |u|^2$ and $|s|^2 + |t|^2 + |u|^2$ approximations over the full calculation.
In the first case, the approximation is good within $\pm10\%$ over the whole range apart in the low invariant-mass region at both low and large rapidity difference.
The low rapidity-difference region possesses remnants of the tri-bosons contributions that have a di-jet invariant mass around the $\PW$-boson mass.
It is therefore expected that the $|t|^2 + |u|^2$ approximation fails in this region.
The second plot, where the $|s|^2 + |t|^2 + |u|^2$ approximation is considered, displays a better behaviour in the previously mentioned region.
The full calculation is approximated at the level of $\pm5\%$ apart in the region where $\Delta y_{\Pj\Pj} < 2$.

>>>>

In Fig.~\ref{fig:ratio2d_LO}, ratios of double-differential cross sections in the variables  $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$ are shown.\footnote{In Fig.~\ref{fig:ratio2d_LO}, the level of the accuracy of the predictions in each bin is around a per mille.}
Two plots are displayed: the ratios of the $|t|^2 + |u|^2$ and $|s|^2 + |t|^2 + |u|^2$ approximations over the full calculation.
In the first case, the approximation is good within $\pm10\%$ over the whole range apart in the low invariant-mass region at both low and large rapidity difference.
The low rapidity-difference region possesses remnants of the tri-bosons contributions that have a di-jet invariant mass around the $\PW$-boson mass.
It is therefore expected that the $|t|^2 + |u|^2$ approximation fails in this region.
The second plot, where the $|s|^2 + |t|^2 + |u|^2$ approximation is considered, displays a better behaviour in the previously mentioned region.
The full calculation is approximated at the level of $\pm5\%$ apart in the region where $|\Delta y_{\Pj\Pj}| < 2$.

### 2018-03-04 Draft/scanNLO.tex
Start of implementing Ansgar's comments

These differences are much more evident in differential distributions.
In Fig.~\ref{fig:mjjdyjj_1d_1}, we show the distributions in the di-jet invariant mass $m_{\Pj\Pj}$ and rapidity separation $|\Delta y_{\Pj\Pj}|$.
For large $m_{\Pj\Pj}$ and large $|\Delta y_{\Pj\Pj}|$, as expected, the VBS approximation is performing well and its $s$-channel extension agree with the full calculation within $10\%$ per cent.
% On the other hand, for low invariant and low rapidity separation the the VBS approximation is failing at a level of at least $30\%$.
% In this very same region, the VBS approximation with $s$-channel is good within $10\%$.
This is in contrast with the region  $200 \GeV < m_{\Pj\Pj} < 500 \GeV$ and $2<|\Delta y_{\Pj\Pj}|<2.5$, where
the difference between the $|t|^2+|u|^2$ approximation and the full computation reaches $30\%$.
The inclusion of $s$-channels cures partly this behaviour in this region.
Still, for the very low $m_{\Pj\Pj}$ a difference of about $5\%$ remains.
This tends to indicate that interference contributions and/or non-factorisable QCD corrections play a non-negligible role in this phase-space region.

>>>>

These differences are much more evident in differential distributions.
In Fig.~\ref{fig:mjjdyjj_1d_1}, we show the distributions in the di-jet invariant mass $m_{\Pj\Pj}$ and rapidity separation $|\Delta y_{\Pj\Pj}|$.
For large $m_{\Pj\Pj}$ and large $|\Delta y_{\Pj\Pj}|$, as expected, the VBS approximation is performing well and its $s$-channel extension agrees with the full calculation within $10\%$ per cent.
% On the other hand, for low invariant and low rapidity separation the the VBS approximation is failing at a level of at least $30\%$.
% In this very same region, the VBS approximation with $s$-channel is good within $10\%$.
This is in contrast with the region  $200 \GeV < m_{\Pj\Pj} < 500 \GeV$ and $2<|\Delta y_{\Pj\Pj}|<2.5$, where
the difference between the $|t|^2+|u|^2$ approximation and the full computation reaches $30\%$.
The inclusion of $s$-channels cures partly this behaviour in this region.
Still, for the very low $m_{\Pj\Pj}$ a difference of about $5\%$ remains.
This tends to indicate that interference contributions and/or non-factorisable QCD corrections play a non-negligible role in this phase-space region.
